KIN: A method to infer relatedness from low-coverage ancient DNA

### 2022-05-14 paper_overleaf/main.tex

We further applied KIN to a genome-wide dataset of 118 ancient individuals from the Lech Valley \cite{mittnik_kinship-based_2019}. We compared our relatedness estimates to those obtained by the authors using READ and lcMLkin. We found that in most cases the three methods agree. Only $<5\%$ of pairs (315 out of 6903) show a different relatedness for KIN and READ (table S2). However, most of these are at very low confidence for both the methods ($\Delta LL>1$ for KIN and $sd<1$ for READ). There are 28 pairs where KIN shows confident results, and when we plot the raw heterozygosity estimates along the genome, in many cases the prediction by KIN is in agreement with the observed proportion of differences along the genome (fig. S\ref{figS8:eg1}). Comparisons with lcMLkin are spare since lcMLkin shows reliable results for $\appr 53\%$ pairs, and we find 16 pairs for which lcMLkin gives a reliable result (\cite{mittnik_kinship-based_2019}), and KIN shows $\Delta LL>$1. We again investigated these comparisons in more detail, and found that at least in cases where there is good amount of overlapping sites ($\geq20,000$), we can clearly see that the proportion of differences along the genome aligns with the prediction of KIN (fig. S\ref{figS9:eg2}).

>>>>

We further applied KIN to a genome-wide dataset of 118 ancient individuals from the Lech Valley \cite{mittnik_kinship-based_2019}. We compared our relatedness estimates to those obtained by the authors using READ and lcMLkin. We found that in most cases the three methods agree. Only $<5\%$ of pairs (315 out of 6903) show a different relatedness for KIN and READ (table S2). However, most of these are at very low confidence for both the methods ($\Delta LL>1$ for KIN and $sd<1$ for READ). There are 28 pairs where KIN shows confident results, and when we plot the raw heterozygosity estimates along the genome, in many cases the prediction by KIN is in agreement with the observed proportion of differences along the genome (fig. S\ref{figS8:eg1}). Comparisons with lcMLkin are sparse since lcMLkin shows reliable results for $\appr 53\%$ pairs, and we find 16 pairs for which lcMLkin gives a reliable result (\cite{mittnik_kinship-based_2019}), and KIN shows $\Delta LL>$1. We again investigated these comparisons in more detail, and found that at least in cases where there is good amount of overlapping sites ($\geq20,000$), we can clearly see that the proportion of differences along the genome aligns with the prediction of KIN (fig. S\ref{figS9:eg2}).
